---
title: "Hierarchical and Mixed Effects Model"
output: html_notebook
---

# Hierarchical and Mixed Effects

Hierarchical model: nested data, not independent
* student test scores not independent from one another if in same class, school, etc.
* also known as nested models, multi-level models
* random effect that pools information across groups
* model with random and fixed effects is mixed effects
* repeated measures or paired tests: same individuals over time

Basic model; mean across all groups is beta
with multiple interecepts; grups 2 and 3 agains global intercept
-if want mean of each group - essentially anova
interactions when groups have diffenrt slopes
lm(y~x-1) estimates an intercept for each x

geom_abline with manually specified slope and intercept works better with compleicated models

```{r}
# First, run a model without considering different intercept for each group
out1 <- lm( response ~ x, multIntDemo)
summary(out1)
plot_output1(out1)

# Considering same slope but different intercepts, the random-effect should be the second on the right hand side
out2 <- lm( response ~ x + group - 1, multIntDemo)
summary(out2)
plot_output2(out2)

# Considering different slope and intercept for each group (i.e., an interaction), the random-effect should be the second on the right hand side
out3 <- lm( response ~ x + group - 1 + x:group, multIntDemo)
summary(out3)
plot_output3(out3)
```


students nested in classes classes nested in schools
nested relationships creates multi level model
mapping from one distribution to another; pool information; outliers have less impact and are treated as random effects
random effect assumes normal distribution

```{r}
library(lme4)

# random effects intercept model
outLmer <- lmer( response ~ x + ( 1 | group), multIntDemo)

# Look at model outputs 
summary( outLmer )
tidy( outLmer )

# Extract predictor variables and plot
extractAndPlotOutput(outLmer)

# Random Effect slopes and intercepts
outLmer2 <- lmer( response ~ ( x | group ), multIntDemo)
summary(outLmer2)
tidy(outLmer2)

# Extract and plot
plotOutput(outLmer2)
```

```{r}
# Mixed effect model
lmerModel <- lmer(mathgain ~ sex + 
                  mathprep + mathknow + (1|classid) +
                  (1|schoolid), data = studentData, na.action = "na.omit",
                  REML = TRUE)
summary(lmerModel)

# Extract and plot 
extractAndPlot(lmerModel)
```

## Linear Mixed Effects Models

Demographic stocasticitiy: small populations
* Random intercept with fixed mean: `(1 | group)`
* Intercepts vary among g1 and g2 within g2: `(1 | g1/g2)`
* Random Intercepts for 2 variables: `(1 | g1) + (1 | g2)`
* correlated random slope and intercept: `x + x( | g)`
* uncorrelated random slope and intercetp: `x + (x || g)`

```{r}
# First, build a lmer with state as a random effect. Then look at the model's summary and the plot of residuals. 
birthRateStateModel <- lmer(BirthRate ~ (1|State), data = countyBirthsData)
summary(birthRateStateModel)
# plot residuals
plot(birthRateStateModel)

# Next, plot the predicted values from the model on top of the plot shown during the video.
countyBirthsData$birthPredictState <-  predict(birthRateStateModel, countyBirthsData)
ggplot() + theme_minimal() +
    geom_point(data =countyBirthsData,
               aes(x = TotalPopulation, y = BirthRate)) + 
        geom_point(data = countyBirthsData,
                   aes(x = TotalPopulation, y = birthPredictState),
                   color = 'blue', alpha = 0.5) 

# Include the AverageAgeofMother as a fixed effect within the lmer and state as a random effect
ageMotherModel <- lmer( BirthRate ~ AverageAgeofMother + (1|State), countyBirthsData)
summary(ageMotherModel)

# Compare the random-effect model to the linear effect model 
summary(lm(BirthRate ~ AverageAgeofMother, data = countyBirthsData))


# Include the AverageAgeofMother as a correlated random-effect slope parameter
ageMotherModelRandomCorrelated <- lmer( BirthRate ~ AverageAgeofMother + (AverageAgeofMother | State),countyBirthsData)
summary(ageMotherModelRandomCorrelated)



# Include the AverageAgeofMother as an uncorrelated random-effect slope parameter
ageMotherModelRandomUncorrelated <- lmer( BirthRate ~ AverageAgeofMother + (AverageAgeofMother || State),countyBirthsData)
summary(ageMotherModelRandomUncorrelated)
```

Age as fixed effect with Age as random effect correlated with Age; State is random effect; 

REML: restricted expected maximum likelihood 
NO P-VALUES

```{r}
# Extract the fixed-effect coefficients
fixef(out)

# Extract the random-effect coefficients
ranef(out)

# Estimate the confidence intervals 
confint(out)


# Extract out the parameter estimates and confidence intervals and manipulate the data
dataPlot <- data.frame(cbind( fixef(out), confint(out)[ 3:4, ]))
rownames(dataPlot)[1] <- "Intercept"
colnames(dataPlot) <- c("mean", "l95", "u95")
dataPlot$parameter <- rownames(dataPlot)

# Print the new dataframe
print(dataPlot)

# Plot the results using ggplot2
ggplot(dataPlot, aes(x = parameter, y = mean,
                     ymin = l95, ymax = u95)) +
    geom_hline( yintercept = 0, color = 'red' ) +
    geom_linerange() + geom_point() + coord_flip() + theme_minimal()
```













```{r}
x <- rio::import("~/github/weather/weather_comp/sun.csv") %>% filter(location == "Charlotte, North Carolina, United States")  %>% 
  mutate(mon_num = as.character(match(month, month.abb))) %>% 
  mutate(date = paste0("2018-",mon_num,"-01")) %>% 
  mutate(date = as.Date(date)) %>% 
  mutate(rana = runif(n(),0,1),
         ranb = runif(n(),0,1),
         ranc = runif(n(),0,1)) %>% 
  filter(data_key == "sun_pct")

ggplot(x, aes(x = date)) + 
  geom_area(aes(y = data_value, fill = data_key)) +
  scale_x_date(labels = scales::date_format("%b"), breaks = scales::date_breaks("months")) +
  scale_fill_manual(values = c("#e0c73a"), name = "Time") +
  geom_col(aes(y = rana), stata = "identity", alpha = 0.9, color = "blue", fill = "blue") +
  guides(fill = FALSE, color = FALSE) +
  theme(
    plot.background = element_rect(fill="lightsteelblue2"),
    panel.background = element_rect(fill="lightsteelblue2"),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.title.x=element_blank(),
    axis.title.y=element_blank())
```


Temperature (red-orange)

```{r}
temp <- rio::import("~/github/weather/weather_comp/temps.csv") %>% filter(location == "Charlotte, North Carolina, United States") %>% 
  mutate(mon_num = as.character(match(month, month.abb))) %>% 
  mutate(date = paste0("2018-",mon_num,"-01")) %>% 
  mutate(date = as.Date(date)) 

yy <- y %>% select(-temp_avg, -mon_num, -month) %>% gather(data_key, data_value, -location, -date)

ggplot(y, aes(x = date)) +
  geom_ribbon(aes(ymax = temp_high_rec, ymin = temp_high_avg), fill = "#871528", alpha = 0.2) +
  geom_ribbon(aes(ymax = temp_high_avg, ymin = temp_low_avg), fill = "#CC211B", alpha = 0.6) +
  geom_ribbon(aes(ymax = temp_low_avg, ymin = temp_low_rec), fill = "#871528", alpha = 0.2) +
  geom_label(data = yy, aes(date, data_value, label = data_value, fill = data_value, color = data_value), label.size = 0)+
  viridis::scale_fill_viridis(option="magma") +
  viridis::scale_color_viridis(option = "viridis", direction = -1) +
  scale_x_date(labels = scales::date_format("%b"), breaks = scales::date_breaks("months")) +
  guides(fill = FALSE, color = FALSE) +
  theme(
    plot.background = element_rect(fill="#3D0116"),
    panel.background = element_rect(fill="#3D0116"),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.title.x=element_blank(),
    axis.title.y=element_blank(),
    legend.key = element_blank(),
    legend.text = element_text(color = "white"),
    axis.text.x = element_text(color = "white"),
    axis.text.y = element_text(color = "white"),
    axis.ticks.x = element_line(color = "white"),
    axis.ticks.y = element_line(color = "white"))
```

https://stackoverflow.com/questions/48424682/how-do-i-limit-the-range-of-the-viridis-colour-scale

percipitation (blue)
```{r}
precip <- rio::import("~/github/weather/weather_comp/precip.csv") %>% filter(location == "Charlotte, North Carolina, United States") %>% 
  mutate(month = factor(month, levels = month.abb))

ggplot(precip, aes(month, data_value, fill=data_key)) +
            geom_bar(stat = "identity", position = "dodge") +
            scale_fill_manual(values = c("#4E7AC7","#ADD5F7"), name = "Type") +
            labs(y = "Precipitation (Inches)", x = "Month") +
            theme(
              text = element_text(color = "white"),
              plot.background = element_rect(fill="#16193B"),
              panel.background = element_rect(fill="#16193B"),
              panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
              axis.title.x=element_blank(),
              axis.title.y=element_blank(),
              legend.background = element_rect(fill="#16193B", size=.5, linetype="dotted"),
              legend.position="bottom",
              legend.key = element_blank(),
              axis.text.x = element_text(color = "white"),
              axis.text.y = element_text(color = "white"),
              axis.ticks.x = element_line(color = "white"),
              axis.ticks.y = element_line(color = "white"))
```



Sun (Yellow)
https://color.adobe.com/Teal-Orange-color-theme-2592429/edit/?copy=true&base=2&rule=Custom&selected=3&name=Copy%20of%20Teal%20%20Orange&mode=rgb&rgbvalues=0,0.180392,0.25098,0.164706,0.341176,0.411765,1,1,1,0.980392,0.741176,0.290196,0.980392,0.588235,0&swatchOrder=0,1,2,3,4

```{r}
sunx <- rio::import("~/github/weather/weather_comp/sun.csv") %>% filter(location == "Charlotte, North Carolina, United States") %>% 
  mutate(date = as.Date(date)) %>% 
  mutate(data_key = factor(data_key, c("pct_night", "cloudy_pct", "sun_actual_pct"))) 
  
ggplot(sunx, aes(x = date)) + 
  geom_area(aes(y = data_value, fill = data_key)) +
  scale_x_date(labels = scales::date_format("%b"), breaks = scales::date_breaks("months")) +
  scale_fill_manual(values = c("#455C7B", "#76696C", "#FFAA5C"), 
                    name = element_blank(),
                    labels=c("Night", "Cloudy", "Sunny")) +
  theme(
    plot.background = element_rect(fill="#AC6C82"),
    panel.background = element_rect(fill="#AC6C82"),
    panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
    axis.title.x=element_blank(),
    axis.title.y=element_blank(),
    legend.background = element_rect(fill="#AC6C82", size=.5, linetype="dotted"),
    legend.position="bottom",
    legend.key = element_blank(),
    legend.text = element_text(color = "white"),
    axis.text.x = element_text(color = "white"),
    axis.text.y = element_text(color = "white"),
    axis.ticks.x = element_line(color = "white"),
    axis.ticks.y = element_line(color = "white"))

```

create niceness index
perfect score: temp = average(high and low) = 76; sunny actual = .5; rain = 0

```{r}
tempc <- rio::import("~/github/weather/weather_comp/temps.csv")

sunc <- rio::import("~/github/weather/weather_comp/sun.csv") %>% spread(data_key, data_value)

precipc <- rio::import("~/github/weather/weather_comp/precip.csv") %>% spread(data_key, data_value)

comfort <- tempc %>% inner_join(sunc) %>% inner_join(precipc) %>% 
  mutate(date = as.Date(paste0("2018-",month,"-01"),"%Y-%b-%d")) %>% 
  mutate(date = replace(date, month == "Dec", as.Date("2018-12-31"))) %>% 
  mutate(avg_temp = (temp_high_avg + temp_low_avg)/2,
         rain_d = rain/30,
         snow_d = snow/30) %>% 
  mutate(temp_comfort = ((temp_high_avg-73)/73) + ((temp_low_avg-60)/60)) %>% 
  mutate(temp_comfort = replace(temp_comfort, temp_comfort < -1, -1)) %>% 
  mutate(other_comfort = cloudy_pct + pct_night + rain_d + (.2 + snow_d)) %>% 
  mutate(other_comfort = replace(other_comfort, other_comfort > 1, 1)) %>% 
  mutate(comfort = if_else(temp_comfort >= 0, temp_comfort + other_comfort, temp_comfort - other_comfort)) %>% 
  select(location, date, comfort) %>% 
  group_by(location) %>% 
  complete(date = seq.Date(min(date), max(date), by="day")) %>% 
  mutate(ci = zoo::na.approx(comfort, na.rm =FALSE)) 

x<- comfort %>% filter(location == "Jerusalem, Israel") 
ggTimeSeries::ggplot_calendar_heatmap(x, 'date', 'ci') +
  scale_fill_gradient2(low = "blue", mid = "green", high = "red",
                             midpoint = 0,
                             na.value = "grey") +
        guides(fill = FALSE, color = FALSE) +
        ggthemes::theme_solarized() %+replace%
        theme(
          panel.border = element_blank(),
          panel.background = element_blank(),
          strip.background = element_blank(),
          strip.text = element_blank(),
          axis.title.x = element_blank(),
          axis.text.x = element_text(size = 14),
          axis.title.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.y = element_blank(),
          axis.line.y = element_blank()
        )
```


```{r}
cluster <- tempc %>% inner_join(sunc) %>% inner_join(precipc) %>% 
      group_by(location) %>% 
      summarise(low_temp = mean(temp_low_avg),
                high_temp = mean(temp_high_avg),
                cloudy = mean(cloudy_pct),
                rain = sum(rain),
                snow = sum(snow)) %>% 
      distinct(location, .keep_all = TRUE) %>% 
      filter(is.na(cloudy)==FALSE) %>% 
      mutate_at(vars(-location), funs(scale)) %>% 
      mutate_at(vars(-location), as.numeric)

# distance <- factoextra::get_dist(scale(cluster[,2:6]))

c2 <- scale(cluster[-1])

k2 <- kmeans(c2, centers = 5, nstart = 25)

k2_loc <- cluster %>% 
    mutate(avg_temp = (low_temp + high_temp)/2) %>% 
    mutate(precip = rain + snow) %>% 
    mutate(precip = replace(precip, precip > 5, 5)) 

k2_loc$cluster <- k2$cluster
k2_loc <- as_tibble(k2_loc)

group.colors <- c("1" = "#d33682", "2" = "#859900", "3" ="#073642", "4" = "#268bd2", "5" = "#cb4b16")

kmeans_plot <- ggplot(data = k2_loc) + 
    geom_point(aes(x = precip, y = avg_temp, color = cluster, text = location)) +
    geom_point(data=k2_loc[k2_loc$location=="Kingston, Jamaica", ], 
               aes(x = precip, y = avg_temp, color = cluster, text = location), 
               size=6, alpha = 0.8, shape = 15) +
    scale_color_manual(values = group.colors) +
    ggthemes::theme_solarized() %+replace%
        theme(
          axis.title.x = element_blank(),
          axis.ticks.x = element_blank(),
          axis.text.x = element_blank(),
          axis.title.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.y = element_blank()
          # axis.line.y = element_blank()
        )


plotly::ggplotly(kmeans_plot, tooltip = "text") 

kmeans_plot %>% ggplotly(tooltip = "text") %>% layout(annotations = a)


gg <- plotly::ggplotly(kmeans_plot)

a <- c(
  autotick = FALSE,
  ticks = "outside",
  tick0 = 0,
  dtick = 0.25,
  ticklen = 5,
  tickwidth = 2
)

gg <- layout(gg, layout(xaxis = a))
gg

# labels = c("Hot and Wet", "Comfortable", "Cold and Wet", "Cold and Dry", "Hot and Dry") 



xstart = 0
xmax = 1
xmin = -1
padding = 0.05
ypos = -0.1

a <- c(
            x=xmin,
            y=ypos,
            ax=xstart + padding,
            ay=ypos,
            xref='x',
            axref='x',
            yref='paper',
            ayref='paper',
            showarrow=TRUE,
            arrowhead=2,
            arrowsize=1,
            arrowwidth=3,
            arrowcolor='#0000ff'
)

pal <- c("#d33682", "#859900", "#073642", "#268bd2", "#cb4b16")

plot_ly(data = k2_loc, 
        x = ~precip, 
        y = ~avg_temp, 
        color = ~cluster,
        colors = pal,
        text = ~location,
        hoverinfo = 'text',
        type = "scatter", 
        mode = "markers") %>% 
    layout(yaxis = list(tile = "Temp -->"), xaxis = list(title = "Precipitation -->"))


```









```{r}
set.seed(666)
cluster <- tempc %>% inner_join(sunc) %>% inner_join(precipc) %>% 
      group_by(location) %>% 
      summarise(low_temp = mean(temp_low_avg),
                high_temp = mean(temp_high_avg),
                cloudy = mean(cloudy_pct),
                rain = sum(rain),
                snow = sum(snow)) %>% 
      distinct(location, .keep_all = TRUE) %>% 
      filter(is.na(cloudy)==FALSE) %>% 
      mutate_at(vars(-location), funs(scale)) %>% 
      mutate_at(vars(-location), as.numeric)

c2 <- scale(cluster[-1])

k2 <- kmeans(c2, centers = 5, nstart = 25)

k2_loc <- cluster %>% 
    mutate(avg_temp = (low_temp + high_temp)/2) %>% 
    mutate(precip = rain + snow) %>% 
    mutate(precip = replace(precip, precip > 5, 5))

k2_loc$cluster <- k2$cluster
k2_loc <- as_tibble(k2_loc) 

k2_loc$cluster = recode(k2_loc$cluster, `4` = "Hot and Wet", 
                 `1` = "Comfortable", 
                 `3` = "Cold and Wet", 
                 `2` = "Cold and Dry", 
                 `5` = "Hot and Dry")

cities <- k2_loc %>% filter(location %in% c("Kingston, Jamaica", "Jerusalem, Israel"))

plot_ly(data = k2_loc, 
        x = ~precip, 
        y = ~avg_temp, 
        color = ~cluster,
        colors = c("#268bd2", "#073642", "#859900", "#d33682", "#cb4b16"),
        text = ~location,
        hoverinfo = 'text',
        type = "scatter", 
        mode = "markers") %>% 
    layout(yaxis = list(title = "Temp -->", 
                        showticklabels = FALSE), 
           xaxis = list(title = "Precipitation -->",
                        showticklabels = FALSE),
           plot_bgcolor = "#fdf6e3",
           paper_bgcolor = "#fdf6e3", 
           annotations = list(x = cities$precip,
                              y = cities$avg_temp,
                              text = cities$location,
                              # bordercolor='#c7c7c7',
                                borderwidth=0,
                                borderpad=2,
                                bgcolor="white",
                                opacity=0.9))
```

